--- 
title: "DfT Introduction to R"
author: "Will Bowditch" #Add names as chapters are added!
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: departmentfortransport/ds-processes
description: "An introduction to working with R at the Department for Transport"
---
```{r include=FALSE}
require(dplyr)

# Load example data for Road Casualties in Great Britain 1969-84
data(Seatbelts)

# Look at the data
sb <- data.frame(Seatbelts)
```

# Why do we need _another_ book?

`R` is a very flexible programming language, which inevitably means there are lots of ways to achieve the same result. This is true of all programming languages, but is particularly exaggerated in `R` which makes use of ['meta-programming'](http://adv-r.had.co.nz/). 

For example, here is how to calculate a new variable using standard R:
```{r}
# sb contains data for Road Casualties in Great Britain 1969-84
sb$Percent_killed_in_rear <- sb$DriversKilled / sb$rear
```

Here's the same thing using Tidyverse style R:
```{r, eval = FALSE}

sb %>%
  dplyr::mutate(
    Percent_killed_in_rear = DriversKilled / rear
  )
```

These coding styles are quite different. As people write more code across the Department, it will become increasingly important that code can be handed over to other R users. It is much easier to pick up code written by others if it uses the same coding style you are familiar with. 

This is the main motivation for this book, to establish a way of coding that represents a sensible default for those who are new to R that is readily transferable across DfT. 


## Coding standards

Related to this, the Data Science team maintain a [coding standards document](https://departmentfortransport.github.io/ds-processes/Coding_standards/r.html), that outlines some best practices when writing R code. This is not prescriptive and goes beyond the scope of this document, but might be useful for managing your R projects. 

## Adding to the book

This book is not static - new chapters can be added and current chapters can be amended. If you want to edit or write a chapter for the book, check out the [GitHub page](https://github.com/departmentfortransport/ds-r-training). 


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

